---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# stride

<!-- badges: start -->
This R package implements nonparametric estimation of survival models for mixture data.
Mixture data is data collected from multiple populations of interest, but we do not know from which population each observation belongs.  The corresponding references are:

Garcia, T.P. and Parast, L. (2020). Dynamic landmark prediction for mixture data. Biostatistics,  doi:10.1093/biostatistics/kxz052.

Garcia, T.P., Marder, K. and Wang, Y. (2017). Statistical modeling of Huntington disease onset.
In Handbook of Clinical Neurology, vol 144, 3rd Series, editors Andrew Feigin and Karen E. Anderson.

Qing, J., Garcia, T.P., Ma, Y., Tang, M.X., Marder, K., and Wang, Y. (2014).
Combining isotonic regression and EM algorithm to predict genetic risk under monotonicity constraint.
Annals of Applied Statistics, 8(2), 1182-1208.

Wang, Y., Garcia, T.P., and Ma. Y. (2012).  Nonparametric estimation for censored mixture data with
application to the Cooperative Huntington's Observational Research Trial. Journal of the American Statistical Association, 107, 1324-1338.
<!-- badges: end -->



## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tpgarcia/stride", force=TRUE)
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example, results="hide", message=FALSE}
library(stride)

# Setup parameters to generate the data
set.seed(1)  ## set the random seed generator.
censoring.rate <- 40  ## set the random censoring rate as 40%
p <- 2  ## number of population subgroups
n <- 2000  ## total sample size
m <- 4  ## number of mixture proportions.

## NOT RUN
# simu.setting <- "HD-No-Covariates"  ## simulate the distribution functions similar to the Huntington disease study in Garcia and Parast (2020) where the distribution functions do NOT depend on covariates. The GenerateData() function still generates covariates, but the simulated distribution functions do not depend on them.

simu.setting <- "HD-With-Covariates"  ## simulate the distribution functions similar to the Huntington disease study in Garcia and Parast (2020) where the distribution functions depend on covariates.

## compute the finite set of mixture proportions. We assume the mixture proportions are computed externally.
qvs <- qvs.values(p,m)

## generate the data
data.gen <- GenerateData(n,p,m,qvs,censoring.rate,simu.setting)
x <- data.gen$x  ## possibly censored survival outcomes
delta <- data.gen$delta  ## censoring indicator, 1= event time observed, 0= event time censored
q <- data.gen$q  ## mixture proportion for each individual
ww <- data.gen$ww  ## observed w-covariate for each individual
zz <- data.gen$zz  ## observed z-covariate for each individual


## Estimation procedures to run to estimate F(t|t0,z,w)
update.qs <- FALSE ## We do not update the mixture proportions q's. Updating mixture proportions is still in a trial phase.
run.NPMLEs <- TRUE  ## We run the type I nonparametric maximum likelihood estimator. 
run.NPNA <- TRUE
run.NPNA_avg <- FALSE
run.NPNA_wrong <- FALSE
run.OLS <- FALSE
run.WLS <- FALSE
run.EFF <- FALSE



## The distribution function we are estimating is F(t|t0,z,w). 
tval <- seq(0,80,by=5)  ## tval refers to "t" in F(t|t0,z,w) 
tval0 <- c(0,20,30,40,50) ##tval0 refers to "t0" in F(t|t0,z,w)
z.use <- c(0,1)  ## z.use refers to "z" in  F(t|t0,z,w)
w.use <- seq(35,55,by=1)  ## w.use refers to "w" in F(t|t0,z,w)

run.prediction.accuracy <- TRUE
do_cross_validation_AUC_BS <- FALSE
know.true.groups <- TRUE
## true group membership (needed to compute the AUC/BS for simulated data)
true.groups <- data.gen$true.groups


## Perform the estimation			
estimators.out <- stride.estimator(n,m,p,qvs,q,
				x,delta,ww,zz,
				run.NPMLEs,
				run.NPNA,
				run.NPNA_avg,
				run.NPNA_wrong,
				tval,tval0,
				z.use,w.use,
				update.qs,
				know.true.groups,
				true.groups,
				run.prediction.accuracy,
				do_cross_validation_AUC_BS)
```
